---
title: "Project2"
author: "yleung"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


## R Libraries
```{r libraries,results='hide', message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(aod)
library(caret)
library(glmnet)
library(corrplot)
library(MASS)
library(regclass)
library(FactoMineR)
library(factoextra)
library(pROC)
library(psych)
library(MLmetrics)
```

##Import Data & ETL
```{r Import-Data-ETL}
xlsx_kobe <- read_excel("C:\\Users\\Yat\\Documents\\MSDS\\MSDS 6372\\Project 2\\project2Data.xlsx")

#Take out an ID column from when data was imported
kobe<-xlsx_kobe[,c(2:29)]

#Re-code the character columns into factors
kobe<-as.data.frame(unclass(kobe))
kobe$shot_made_flag<-as.factor(kobe$shot_made_flag)
kobe$shot_type<-as.factor(kobe$shot_type)
kobe$shot_zone_area<-as.factor(kobe$shot_zone_area)
kobe$shot_zone_basic<-as.factor(kobe$shot_zone_basic)
kobe$shot_zone_range<-as.factor(kobe$shot_zone_range)
kobe$game_date<-as.factor(kobe$game_date)
kobe$season<-as.factor(kobe$season)
kobe$period<-as.factor(kobe$period) #change period into factor?
kobe$playoffs<-as.factor(kobe$playoffs)

#subsetting variables (sans IDs)
kobe<-kobe[,c(1:2,5:20,22:25,27:28)]
kobe<-na.omit(kobe)


#data check
#head(kobe)
summary(kobe)
str(kobe)
```


```{r prin-comp}
#numerical variables
kobe.NV<-kobe[,c(3:7,11:12,23:24)]
kobe.NV<-na.omit(kobe.NV)
fit<-prcomp(~., data=kobe.NV, cor=TRUE)
summary(fit) # print variance accounted for 
loadings(fit) # pc loadings 
plot(fit,type="lines") # scree plot 
fit$scores # the principal components
biplot(fit,expand=10, xlim=c(-0.15, 0.05), ylim=c(-0.1, 0.05))

# Varimax Rotated Principal Components
# Extract, rotate and retain 5 PCs 
component.retained <- principal(kobe.NV, nfactors=5, rotate="varimax")
component.retained


# Principal Axis Factor Analysis
axis.fit <- factor.pa(kobe.NV, 5)
axis.fit

#PCA
res.pca1 <- prcomp(kobe.NV, scale = TRUE)
fviz_eig(res.pca1)

fviz_pca_var(res.pca1,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )

#numerical variables with lat removed
kobe.NV2<-kobe[,c(4:7,11:12,23:24)]
kobe.NV2<-na.omit(kobe.NV2)
fit2<-prcomp(~., data=kobe.NV2, cor=TRUE)
summary(fit2) # print variance accounted for 
loadings(fit2) # pc loadings 
plot(fit2,type="lines") # scree plot 
fit2$scores # the principal components
biplot(fit2,expand=10, xlim=c(-0.15, 0.05), ylim=c(-0.1, 0.05))

# Varimax Rotated Principal Components
# Extract, rotate and retain 5 PCs 
component.retained2 <- principal(kobe.NV2, nfactors=5, rotate="varimax")
component.retained2


# Principal Axis Factor Analysis
axis.fit2 <- factor.pa(kobe.NV2, 5)
axis.fit2

#PCA
res.pca2 <- prcomp(kobe.NV2, scale = TRUE)
fviz_eig(res.pca2)

fviz_pca_var(res.pca2,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
```

#Contingency Tables
```{r contingency-tbls}
#contingency table samples. coordination variables not included due to large computation time.
list.of.xtabs <- lapply(kobe[,c(7,11:12)], function(x) xtabs(~ x + kobe$shot_made_flag))
list.of.xtabs
```


#Chi-sq test for variables independence
null: the 2 categorical variables are independent.
```{r chi-sq_indep}
#action_type and combined_shot_type
tbl.1<-table(kobe$action_type, kobe$combined_shot_type)
chisq.test(tbl.1) 

#shot_zone_area and shot_zone_range
tbl.2<-table(kobe$shot_zone_area, kobe$shot_zone_range)
chisq.test(tbl.2) 

#shot_zone_area and shot_zone_basic
tbl.3<-table(kobe$shot_zone_area, kobe$shot_zone_basic)
chisq.test(tbl.3) 

#shot_zone_range and shot_zone_basic
tbl.4<-table(kobe$shot_zone_range, kobe$shot_zone_basic)
chisq.test(tbl.4) 
```




##logistic model
```{r base-model}
#shot_made_flag to distance, playoffs and coordinations
glm.fit<-glm(shot_made_flag~loc_x+loc_y+lon+playoffs+shot_distance+shot_zone_area+shot_zone_basic+shot_zone_range,family=binomial(link="logit"),data=kobe)

summary(glm.fit)

```

VIF
Not very useful in logistic regression, typically notr performed for mixed data set with categorical variables.

##Model Selection(s): Stepwise 
Both stepwise models resulted in the same AIC and predictors. Since the question stated that playoffs is a factor in him making the shot, therefore the prediction should use the base.logit (not the two models shown below)
```{r stepwise}
#Stepwise based on the first logit regression
stepwise<- stepAIC(glm.fit,direction="both",trace = FALSE)
summary(stepwise)

```

##Odds Ratio
```{r odds-ratio}

#odds ratio calculation using library(epiDisplay)
#based on the first logit function
exp(cbind("Odds ratio" = coef(stepwise), confint.default(stepwise, level = 0.95)))   

```

LDA is not performed for logistic regression model.


##Prediction
```{r Prediction}
#load the project2Pred.xlsx
pred_kobe <- read_excel("C:\\Users\\Yat\\Documents\\MSDS\\MSDS 6372\\Project 2\\project2Pred.xlsx")
#Take out an ID column from when data was imported
pred_kobe<-pred_kobe[,c(2:29)]

#Re-code the character columns into factors
pred_kobe<-as.data.frame(unclass(pred_kobe))
pred_kobe$shot_made_flag<-as.factor(pred_kobe$shot_made_flag)
pred_kobe$shot_type<-as.factor(pred_kobe$shot_type)
pred_kobe$shot_zone_area<-as.factor(pred_kobe$shot_zone_area)
pred_kobe$shot_zone_basic<-as.factor(pred_kobe$shot_zone_basic)
pred_kobe$shot_zone_range<-as.factor(pred_kobe$shot_zone_range)
pred_kobe$game_date<-as.factor(pred_kobe$game_date)
pred_kobe$season<-as.factor(pred_kobe$season)
pred_kobe$period<-as.factor(pred_kobe$period) #change period into factor?
pred_kobe$playoffs<-as.factor(pred_kobe$playoffs)

#subsetting variables (sans IDs)
pred_kobe<-pred_kobe[,c(1:2,5:20,22:25,27:28)]
pred_kobe<-na.omit(pred_kobe)


#data check
head(pred_kobe)
str(pred_kobe)

#dataframe for the shot_made_flag column in the pred_kobe
pred_shot_made_flag<-pred_kobe

#what library to use for AUC, Mis-Classification Rate, Sensitivity, Specificity and objective / loss function
predictions <- predict(stepwise, pred_shot_made_flag, type="response")
predictions<-as.data.frame(ifelse(predictions>0.5,"1","0"))
colnames(predictions) <- "Predicted shot_made_flag"



#actual and predicted values from model
threshold=0.5
predicted_values<-ifelse(predict(stepwise,type="response")>threshold,1,0)

actual_values<-stepwise$y


#confusion matrix using the training set
conf_matrix<-table(predicted_values,actual_values)
conf_matrix

#Sensitivity of the model
sensitivity(conf_matrix)
#specificity of the model
specificity(conf_matrix)

predicted_prob<-predict(stepwise,type="response")
roccurve <- roc(actual_values, predicted_prob)
plot(roccurve)

#AUC 
auc(roccurve)

#Log Loss Function
LogLoss(predicted_values,actual_values)
```





